### 2024-12-10 README.Rmd
update readme example

First download a friction surface — here using the motorised travel time from
Weiss *et al.* 2020.
```{r }
library(traveltime)
library(terra)

@@

Let's have a look at that
```{r}
plot(friction_surface)
```

Prepare points we would like to calculate travel time from
```{r}
from_here <- tibble::tibble(
  x = c(111.2, 111.9),
  y = c(0.2, 0.35)
)
from_here
```

And calculate the travel time
```{r}
travel_time <- calculate_travel_time(
  friction_surface = friction_surface,
  points = from_here
)
travel_time
```

Et voila!
```{r}
plot(travel_time)
points(from_here, pch = 19)
```

## Let's go to Singapore

Here it is:
```{r}
#install.packages("geodata")
library(geodata)
sin <- gadm(
  country = "Singapore",
  level = 0,
  path = tempdir(),
  resolution = 2
)
plot(sin)
```

We're going to see how long it takes to walk home from an MRT or LRT station. So
we'll download the walking-only friction surface this time.

We can feed in our `sin` `SpatVector` directly as the `extent`, instead of
specifying by hand as above. 
```{r }
library(traveltime)
library(terra)

@@

And where are the stations? The the `stations` data set contains the latitude
and longitude of all LRT and MRT station exits [^1].
```{r}
head(stations)
plot(stations)
```
Let's look at those. Below we plot the friction surface raster, with the vector
boundary as a dashed grey line, and the station exits as grey points.
```{r}
library(tidyterra)
library(ggplot2)

@@

OK, now we want to calculate the walking travel time in minutes across the 
friction surface from the nearest station exit:
```{r}
travel_time_sin <- calculate_travel_time(
  friction_surface = friction_singapore,
  points = stations
)
travel_time_sin
```

>>>>

First download a friction surface --— here we are using the motorised travel
time from Weiss *et al.* 2020. We use the function
`traveltime::get_friction_surface`, specify the surface (type) as `"motor2020"`,
 and provide the spatial extent of interest:
```{r }
library(traveltime)
library(terra)

@@

Let's have a look at that `SpatRaster`:
```{r}
plot(friction_surface)
```

Now, prepare points that we would like to calculate travel time from:
```{r}
from_here <- tibble::tibble(
  x = c(111.2, 111.9),
  y = c(0.2, 0.35)
)
from_here
```

And calculate the travel time from our points `from_here` over the friction 
surface `friction_surface` using the function
`traveltime::calculate_travel_time`:
```{r}
travel_time <- calculate_travel_time(
  friction_surface = friction_surface,
  points = from_here
)
travel_time
```

Et voila! Here is the motorised travel time in minutes for each cell, with our
points in pink.
```{r}
plot(travel_time)
points(from_here, pch = 19, col = "hotpink")
```

## A more tangible example: Walking in Singapore

Let's create a map of the walking time across the island of Singapore from the nearest
[MRT or LRT](https://en.wikipedia.org/wiki/Transport_in_Singapore#Rail_transport)
station.

To do this, we need:

- a map of Singapore
- locations of the stations

Here's our basemap via `geodata`:
```{r}
#install.packages("geodata")
library(geodata)
sin <- gadm(
  country = "Singapore",
  level = 0,
  path = tempdir(),
  resolution = 2
)
plot(sin)
```

We're going to see how long it takes to walk home from a station, so we'll
download the walking-only friction surface this time by specifying 
`surface = "walk2020`.

We can also pass in our basemap `sin`, a `SpatVector`, directly as the `extent`,
instead of specifying by hand as above. We're also only interested in walking 
*on land* so we mask out areas outside of `sin`, that are within the extent of 
the raster:

```{r }
library(traveltime)
library(terra)

@@

The the `stations` data set in this package contains the longitude and latitude
of all LRT and MRT station exits in Singapore[^1].

```{r}
head(stations)
```

Let's look at our data now.

Below we plot the friction surface raster `friction_singapore`, with the vector
boundary of `sin` as a dashed grey line, and `stations` as grey points:

```{r}
library(tidyterra)
library(ggplot2)

@@

OK, now we want to calculate the walking travel time in minutes across the 
friction surface from the nearest station exit:

```{r}
travel_time_sin <- calculate_travel_time(
  friction_surface = friction_singapore,
  points = stations
)
travel_time_sin
```
